Transferability of clinical prediction models for early trauma care

### 2019-06-24 ManuscriptMarkdown.Rmd
Update ManuscriptMarkdown.Rmd

  **Results:** `r total.number.of.patients ` patients were included. Model
  validation median mistriage rate with 95% UI ranged from `r
  lowest.val.mistriage` to `r highest.val.mistriage`, and transferred model
  mistriage rate ranged from `r lowest.tran.mistriage` to `r
  highest.tran.mistriage` across all samples. Worst performance after transfer
  compared to receiving sample validation performance was found in the `r
  colnames(highest.comp.mistriage)` sample with a median mistriage difference of
  `r highest.comp.mistriage`.

  **Conclusions:** Model transfer from the `r colnames(highest.comp.mistriage)`
  to the other sample in the data set lead to an increased mistriage
  rate. Mainly due to a marked increase in overtriage. In all samples, model
  transfer was associated with high levels of uncertainty.

@@

# Results
We analysed data from `r total.number.of.patients ` trauma patients to
investigate the effects of transfer on clinical prediction models (Table 1). The
total number of missing observations across all variables was `r
total.number.of.na ` in the entire study cohort. The sample with the highest
percentage of missing observations was the `r highest.percent.missing.name `
sample with `r highest.percent.missing.number `% incomplete observations. The
variable with the highest number of missing values was `r name.variable.na `,
with `r number.variable.na ` missing values, which constituted `r
round(percent.variable.na) `% of the total values for this variable. The
percentages of missing values for the other model predictors, GCS and SBP was `r
round(gcs.na$percentage)`% and `r round(sbp.na$percentage)`% respectively.

**Table 1. Sample characteristics of multiple imputed data.** Data is presented as medians with interquartile range (IQR) or counts with % as applicable.
```{r, echo = FALSE, results = "asis"} 
## Table 1
options(knitr.kable.NA = '')
kable(table_one_high_low, format = "markdown")
kable(table_one_metro_non_metro, format = "markdown")
kable(table_one_multi_single, format = "markdown")
``` 
Abbreviations: GCS glasgow coma scale, SBP systolic blood pressure, RR respiratory rate, ISS injury severity score, NISS new injury severity score

@@

# Discussion
This study aimed to assess how transfers of clinical prediction models for early
trauma care between different contexts within a single health system affect
mistriage rates. The most notable effects on model performance following model
transfer were seen after transferring the `r colnames(highest.comp.mistriage)`
model. The mistriage rate increased by `r substr(format(highest.comp.mistriage),
1, 5)` when using the `r colnames(highest.comp.mistriage)` model in the other
sample in the same data set. Mainly contributing to this increase in mistriage
was the marked increase in overtriage when performing this model transfer. For
this transfer the 95% UI ranged from `r UI.top.bot[1]` to `r UI.top.bot[2]`,
indicating that our findings are compatible with a `r UI.top.bot[1]` decrease to
a `r UI.top.bot[2]` increase in mistriage. The negative 95% UI of `r
UI.top.bot[1]` would signifiy an improvement in model performance. The higher
95% UI of `r UI.top.bot[2]` would mean a marked increase in model mistriage,
compared to the point estimate of `r substr(format(highest.comp.mistriage), 1,
5)`.

@@

## Conclusion
Transfer of the `r colnames(highest.comp.mistriage)` model lead to an increased
mistriage rate of `r highest.comp.mistriage`. This observed effect was
associated with considerable uncertainty, meaning mistriage might slightly
decrease, but would more likely increase, potentially substantially. In many
other samples, model transfer led to large levels of uncertainty, and in all
samples 95% UI included 0. In conclusion, model transfer is associated with
varying degrees of uncertainty but can potentially lead to increased mistriage,
resulting in increased patient mortality and a misappropriation of resources,
due to increased undertriage and overtriage, respectively. Recognizing the
limitations of this study, the authors believe further studies are warranted due
to the potential economic costs and patient consequences of poor model
transfers.

>>>>

  **Results:** `r total.number.of.patients ` patients were included. Model
  validation median mistriage rate with 95% UI ranged from `r lowest.val.mistriage` 
  to `r highest.val.mistriage`, and transferred model
  mistriage rate ranged from `r lowest.tran.mistriage` to `r highest.tran.mistriage` 
  across all samples. Transfer performance when compared to receiving sample validation 
  performance resulted in subtantial uncertainty in all samples. Worst comparison 
  performance after transfer was found in the `r colnames(highest.comp.mistriage)` 
  sample with a median mistriage difference of `r highest.comp.mistriage`.

  **Conclusions:** In all samples, model transfer was associated with high 
  levels of uncertainty. Notably, model transfer from the `r colnames(highest.comp.mistriage)` 
  sample to the other sample in the data set lead to an increased mistriage
  rate, mainly due to a marked increase in overtriage.

@@

# Results
We analysed data from `r total.number.of.patients ` trauma patients to
investigate the effects of transfer on clinical prediction models (Table 1). The
total number of missing observations across all variables was `r total.number.of.na ` 
in the entire study cohort. The sample with the highest
percentage of missing observations was the `r highest.percent.missing.name `
sample with `r highest.percent.missing.number `% incomplete observations. The
variable with the highest number of missing values was `r name.variable.na `,
with `r number.variable.na ` missing values, which constituted `r round(percent.variable.na) `
% of the total values for this variable. The percentages 
of missing values for the other model predictors, GCS and SBP was `r round(gcs.na$percentage)`% 
and `r round(sbp.na$percentage)`% respectively.

**Table 1. Sample characteristics of multiple imputed data.** Data is presented as medians with interquartile range (IQR) or counts with % as applicable.
```{r, echo = FALSE, results = "asis"} 
## Table 1
options(knitr.kable.NA = '')
# kable(results$table.one.list$Characteristics_table_of_high.volume.vs.low.volume , format = "markdown")
# kable(results$table.one.list$Characteristics_table_of_metropolitan.vs.non.metropolitan , format = "markdown")
# kable(results$table.one.list$Characteristics_table_of_multi.centre.vs.single.centre , format = "markdown")

kable(table_one_high_low, format = "markdown")
kable(table_one_metro_non_metro, format = "markdown")
kable(table_one_multi_single, format = "markdown")
``` 
Abbreviations: GCS glasgow coma scale, SBP systolic blood pressure, RR respiratory rate, ISS injury severity score, NISS new injury severity score

@@

# Discussion
This study aimed to assess how transfers of clinical prediction models for early
trauma care between different contexts within a single health system affect
mistriage rates. In all data sets, model transfer between samples resulted in 
substantial uncertainty, with some variation between different data sets. Comparing 
transferred performance with the reciveing sample validation performance resulted in 
further uncertainty.

The most notable effects on model performance following model
transfer were seen after transferring the `r colnames(highest.comp.mistriage)`
model. The mistriage rate increased by `r substr(format(highest.comp.mistriage), 1, 5)` 
when using the `r colnames(highest.comp.mistriage)` model in the other
sample in the same data set. Mainly contributing to this increase in mistriage
was the marked increase in overtriage when performing this model transfer. For
this transfer the 95% UI ranged from `r UI.top.bot[1]` to `r UI.top.bot[2]`, indicating 
a high level of uncertainty and that our findings are compatible with a `r UI.top.bot[1]` 
decrease to a `r UI.top.bot[2]` increase in mistriage. The negative 95% UI of `r UI.top.bot[1]` 
would signifiy an improvement in model performance. The higher
95% UI of `r UI.top.bot[2]` would mean a marked increase in model mistriage,
compared to the point estimate of `r substr(format(highest.comp.mistriage), 1, 5)`.

@@

## Conclusion
Model transfer resulted in large and varying levels of uncertainty, and in all 
samples 95% UI included 0. Model transfer can therefore be unpredictable, but can 
potentially lead to increased mistriage, resulting in increased patient mortality 
and a misappropriation of resources, due to increased undertriage and overtriage, 
respectively. Transfer of the `r colnames(highest.comp.mistriage)` model lead to an 
increased mistriage rate of `r highest.comp.mistriage`. This observed effect was also
associated with considerable uncertainty, meaning mistriage might slightly
decrease, but would more likely increase, potentially substantially. Recognizing 
the limitations of this study, the authors believe further studies are warranted due
to the potential economic costs and patient consequences of poor model
transfers.
